---
title: "PA1_template"
author: "Yan"
date: "Saturday, September 19, 2015"
output: html_document
---

#Loading and preprocessing the data
Loading data
```{r}
aData <- read.csv("activity.csv",sep = ",", colClasses = c("integer", "Date", "factor"))
head(aData)
```

Process/transform the data 
```{r}
aData$month<- as.numeric(format(aData$date, "%m"))
head(aData)
noNA.Data <- na.omit(aData)
rownames(noNA.Data) <- 1:nrow(noNA.Data)
head(noNA.Data)
```


#What is mean total number of steps taken per day?
##Calculate the total number of steps taken per day

Make a histogram of the total number of steps taken each day
```{r}
library(ggplot2)
ggplot(noNA.Data, aes(date, steps)) + geom_bar(stat = "identity", colour = "darkgreen", fill = "darkgreen", width = 0.5) + facet_grid(. ~ month, scales = "free") + labs(title = "Histogram of Total Number of Steps Taken Each Day", x = "Date", y = "Total number of steps")

```

##Calculate and report the mean and median total number of steps taken per day
Mean total number of steps taken per day:
```{r}
totalSteps <- aggregate(noNA.Data$steps, list(Date = noNA.Data$date), FUN = "sum")$x
mean(totalSteps)

```
Median total number of steps taken per day:
```{r}
median(totalSteps)
```

#What is the average daily activity pattern?
Make a time series plot (i.e. type = "1") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis) 
```{r}
avgSteps <- aggregate(noNA.Data$steps, list(interval = as.numeric(as.character(noNA.Data$interval))), FUN = "mean")
names(avgSteps)[2] <- "meanSteps"

plot(avgSteps$interval, avgSteps$meanSteps, type="l",main = "Time Series Plot of the 5 minute Interval", xlab = "5minute intervals", ylab = "Average number of steps taken", ,col="darkgreen", lwd = 3 )
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
avgSteps[avgSteps$meanSteps == max(avgSteps$meanSteps), ]
```

#Imputing missing values
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(aData))

```
All of the missing values are filled in with mean value for that 5-minute interval 
```{r}
newData <- aData 
for (i in 1:nrow(newData)) {
    if (is.na(newData$steps[i])) {
        newData$steps[i] <- avgSteps[which(newData$interval[i] == avgSteps$interval), ]$meanSteps
    }
}

head(newData)
sum(is.na(newData))
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r}
ggplot(newData, aes(date, steps)) + geom_bar(stat = "identity",
                                             colour = "darkgreen",
                                             fill = "darkgreen",
                                             width = 0.5) + facet_grid(. ~ month, scales = "free") + labs(title = "Histogram of Total Number of Steps Taken Each Day (no missing data)", x = "Date", y = "Total number of steps")
```

The values are slightly different from the estimates of the first part of the assignment. Missing data make no significant to this study.


#Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day
```{r}
newData <- aData
head(newData)
newData$weekdays <- factor(format(newData$date, "%A"))
levels(newData$weekdays)
head(newData)
levels(newData$weekdays) <- list(weekday = c("Monday", "Tuesday",
                                             "Wednesday", 
                                             "Thursday", "Friday"),
                                 weekend = c("Saturday", "Sunday"))
head(newData)
levels(newData$weekdays)
```


Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)


```{r}


avgSteps <- aggregate(steps ~ interval + weekdays, newData, mean)

names(avgSteps)[3] <- "meanSteps"
library(lattice)
xyplot(avgSteps$meanSteps ~ avgSteps$interval | avgSteps$weekdays, 
       layout = c(1, 2), type = "l", 
       xlab = "Interval", ylab = "Number of steps")


```

